\textbf{Designing a Virtual Security Layer for Cloud Content}

### 2015-10-27 tex/paper.tex
Added dirty presentation folder

\clearpage
\pagenumbering{arabic}
\setcounter{page}{1}
\section{Introduction}
Developing larger projects containing static content usually involves using a Content Distribution Network 
to be able to scale to a large user base. The commercial Content Distribution Networks are usually fairly easy to use, 
the content that is to be used in a project is usually simply uploaded and then directly available to the public. 
For secret content this can be a problem and that is what this thesis is about. This work examines ways of enforcing 
access control on content and groups of content in the form of snapshots. A system was developed to make the underlying 
theory work in practice.

\newpage
\section{Related Terminology}
\subsection{Technologies}
\subsubsection{React}
React is a JavaScript library for building user interfaces. React uses both its own virtual DOM and the browser's, 
this makes it able to efficiently update dynamic web pages after a change of state through comparing the old virtual 
DOM with the resulting virtual DOM after the state change and then only update the browser's DOM according to the 
difference between the virtual DOMs~\cite{REACT}.

\subsubsection{Flux}

@@

\subsubsection{REST}
\subsubsection{MongoDB}
    TODO: Write down related terminology, if any
\subsection{Abbreviations}

@@

Copy-on-Write is used in virtual memory management systems~\cite{VIRTCOW}, snapshot functionality in both file 
systems~\cite{FSCOW} and logical volume management systems~\cite{LVMCOW}, and as an optimisation technique for objects and 
types in programming languages\cite{LANGCOW}.

Its principle is that when processes share data in between each other, the data is not copied until one of the processes 
does changes to it. This is an optimisation as the processes does not have to send all of the related data that is in memory, 
rather they only have to send pointers to the data. After many Copy-on-Write's a complex structure can be built up, 
but it is possible to solve that structure~\cite{COPYONWRITE2}.

@@

\subsection{Problem description}
As the current system does not offer proper security measurements, is lacking a lot of features that is needed and does not 
scale very well, a new system should be developed. This work is about examining a way of implementing Copy-on-Write in a 
high level system, which should solve the scalability problem and makes it possible to implement the wanted features, like 
snapshots and concurrent modifications of content.

What to write about
Battle Binary
Security features of CDN's

What is needed
* Security layers
* snapshots
* Virtual file structure
* Versioning of content
* Multi project support
* Auth and audit logs
* Users
\newpage 
\section{Model}
\subsection{Entities}
%\subsubsection*{snapshot} for not having section numbers
\begin{figure}[htp] \centering{
    \includegraphics[scale=0.8]{relation.pdf}}
    \caption{Entity Relationships}
    \label{fig:relation}
\end{figure}

\subsubsection{Content}
Content is an asset contained in a project and/or snapshot, it is a form of virtual file.
It could be described as a link between the snapshot or project and the real file. 
The content can be for example an image, video or binary blob, combined with meta-data.

\subsubsection{Project}
A project is what is created to contain all content related to a real project. Files can be changed within a project and 
the system can contain several projects and their virtual content are completely disjoint.
snapshot
\subsubsection{Snapshot}
A snapshot is a read-only container from the state which the container the was in when the snapshot was created. A snapshot can 
have content which either is automatically dependant on the container which it was created from and updated when that container 
is updated or stay in the state which it was in when the snapshot was created. This results in that the content within a snapshot 
can not be updated independently from the container which it was created from.

\subsubsection{Container}
A container is a virtual folder within the projects which contains content and other containers.

\subsubsection{File}
A file refers to an actual physical file in the file system.

@@

\subsection{JPF}
Java path finder was used to show that the model and plan of how to build the system was sound. The model was built in Java 
with the objective of being as reduced and simple as possible, without loosing any of the cases that needs to be covered 
by the model checker. As the users are mainly going to be handled by external systems they were not included in the model.

Each collection in the persistent storage was emulated by using the built-in ConcurrentHashMap type. 
Each client was represented by a thread and each action taken by the client was randomised. The id hashes which MongoDB is 
using for each entity was imported from the mongo-java-driver-2.13.3 and each object had its own id, generated in the same 
fashion as the real implementation is using, randomly generated by the ObjectId class to minimise collisions that is. 
Further more no locking or transactions were used and the threads were running fully concurrently, without any sleep statements.

@@

\begin{lstlisting}[label=JPFRESULT,caption=Results of JPF run]
elapsed time:       14:26:53
states:             new=160853259,
                    visited=451102505,
                    backtracked=611955764,
                    end=21640
search:             maxDepth=380,
                    constraints=0
choice generators:  thread=160853255 
                    (signal=0,
                    lock=3603938,
                    sharedRef=146989208,
                    threadApi=3,
                    reschedule=10260106), 
                    data=0

@@

\newpage 
\section{Copy-on-Write}
As the persistent storage does not implement transactions or locks a lot of different problems can occur when several 
clients are working on the same data set at the same time. Such problems could be race conditions,

@@

\section{Resulting system}
\subsection{Persistent storage}
\subsubsection{MongoDB}
MongoDB was chosen as the persistent storage because of its internal storage format called BSON, which is very similar to JSON 
which the API is using. As the formats are similar, the process of marshalling and unmarshalling becomes quite easy between 
the core code, MongoDB instance and REST interface. The second reason was that if the system needs to scale in the future it 
is very easy to distribute MongoDB and if needed the system can easily be migrated to Reactive Mongo, which is asynchronous and 
non-blocking and can therefore scale even further~\cite{REACTIVEMONGO}.

\subsection{API}
REST was chosen as the BSON format which is used in MongoDB is almost identical ~\cite{BSON} to the standardised JSON 
format which is used by RESTful services~\cite{JSON}.

\subsubsection{REST Endpoints}
For the frontend to communicate with the backend REST is used, the following endpoints were configured:

\begin{itemize}
  \item projects
      \subitem GET - list all projects
      \subitem POST - create new project
  \item projects/\{id\}
      \subitem GET - get specific project
      \subitem PUT - update existing project
      \subitem DELETE - delete existing project
  \item projects/\{id\}/content
      \subitem GET - list all content in a specific project
      \subitem POST - create new content in a specific project 
  \item projects/\{id\}/content/\{id\}
      \subitem GET - get specific content in a specific project
      \subitem PUT - update existing content in a specific project
      \subitem DELETE - delete existing content in a specific project

  \item projects/\{id\}/snapshots
      \subitem GET - list all snapshots in a specific project
      \subitem POST - create new snapshot in a specific project 
  \item projects/\{id\}/snapshots/\{id\}
      \subitem GET - get specific snapshot in a specific project
      \subitem PUT - update existing snapshot
      \subitem DELETE - delete existing snapshot
  \item projects/\{id\}/snapshots/\{id\}/content
      \subitem GET - list all content in a specific snapshot
      \subitem POST - create new content in a specific snapshot 
  \item projects/\{id\}/snapshots/\{id\}/content/\{id\}
      \subitem GET - get specific content in a specific snapshot
      \subitem PUT - update existing content in a specific snapshot
      \subitem DELETE - delete existing content in a specific snapshot

  \item projects/\{id\}/containers
      \subitem GET - list all containers in a specific project
      \subitem POST - create new container in a specific project 
  \item projects/\{id\}/containers/\{id\}
      \subitem GET - get specific container in a specific project
      \subitem PUT - update existing container
      \subitem DELETE - delete existing container
  \item projects/\{id\}/containers/\{id\}/content
      \subitem GET - list all content in a specific container
      \subitem POST - create new content in a specific container 
  \item projects/\{id\}/containers/\{id\}/content/\{id\}
      \subitem GET - get specific content in a specific container
      \subitem PUT - update existing content in a specific container
      \subitem DELETE - delete existing content in a specific container
\end{itemize}

@@

Stuff to write about:
Modular design, every piece should be interchangable
LDAP - why it was used as standard AUTH

>>>>

\clearpage
\pagenumbering{arabic}
\setcounter{page}{1}
\section{Introduction}
Developing larger projects containing static content usually involves using a Content Distribution Network 
to be able to scale to a large user base. The commercial Content Distribution Networks are usually fairly easy to use, 
the content that is to be used in a project is usually simply uploaded and then directly available to the public. 
For secret content this can be a problem and that is what this thesis is about. This work examines ways of enforcing 
access control on content and groups of content in the form of folders and snapshots. A system was developed to make 
the underlying theory work in practice.

\newpage
\section{Related Terminology}
\subsection{Technologies}
\subsubsection{React}
React is a JavaScript library for building user interfaces. React uses both its own virtual DOM and the browser's, 
this makes it able to efficiently update dynamic web pages after a change of state through comparing the old virtual 
DOM with the resulting virtual DOM after the state change and then only update the browser's DOM according to the 
delta between the virtual DOMs~\cite{REACT}. React can be seen as the system for handling views in front-ends 
implementing a MVC (Model-View-Controller) architecture.

\subsubsection{Reflux}
Explain difference to MVC etc.

@@

\subsubsection{REST}
REST stands for representational state transfer, it is an architectural idea for writing stateless services. These 
services usually use URIs to identify specific resources and HTTP to modify or query these resources~\cite{REST}. 
In this work we use a RESTful API, which means that it is an API following the REST architecture.

\subsubsection{MongoDB}
MongoDB is a document-oriented database which means that it does not have the concept of rows as normal relational 
databases has. Instead each entity in the database is stored as a document which is not fixed to a predefined table 
structure~\cite{MONGODB}. MongoDB lacks support for joins too improve its possibility to scale, which can be a big 
down side to some applications containing the need for such logic.

\subsection{Abbreviations}

@@

Copy-on-Write is used in virtual memory management systems~\cite{VIRTCOW}, snapshot functionality in both file 
systems~\cite{BTRFS} and logical volume management systems~\cite{LVMCOW}, and as an optimisation technique for objects and 
types in programming languages\cite{LANGCOW}.

Its principle is that when processes or nodes share data in between each other, the data is not copied until one of the 
processes does changes to it. This is an optimisation as the processes does not have to send or copy all of the related 
data that is in memory, rather they only have to send pointers to the data. After many Copy-on-Write's a complex structure 
can be built up, but it is possible to solve that structure~\cite{COPYONWRITE2}.

TODO: Polish and extend

@@

\subsection{Problem description}
As the current system does not offer proper security measurements, is lacking a lot of features that is needed and does not 
scale very well, a new system should be developed. This work is about examining a way of implementing Copy-on-Write in a 
high level system like this, which should solve the scalability problem and make it possible to implement wanted features
like snapshots, cloning and concurrent modifications of content.

%What to write about
%Battle Binary
%Security features of CDN's
%What is needed by the logic, like what happens after two consecutive deletes
%
%What is needed
%* Security layers
%* snapshots
%* Virtual file structure
%* Versioning of content
%* Multi project support
%* Auth and audit logs
%* Users

%Stuff to write about:
%Modular design, every piece should be interchangeable
%LDAP - why it was used as standard AUTH

%Using copy-on-write as the sole update strategy has pros and cons. The upside is
%that it is simple to guarantee operation atomicity, and data-structure integrity. The
%downside is that performance relies on the ability to maintain large extents of free
%contiguous disk areas. In addition, random updates to a f i le tend to fragment it,
%destroying sequentiality. A good defragmentation algorithm is required; this is
%described in Section 5

\newpage 
\section{Model}
\subsection{Entities}
%\subsubsection*{snapshot} for not having section numbers
\begin{figure}[htp] \centering{
    \includegraphics[scale=0.8]{relation.pdf}}
    \caption{Entity Relationships}
    \label{fig:relation}
\end{figure}

\subsubsection{Content}
Content is meta data about a file and is stored in a container, it is a form of virtual file.
The content can refer to for example an image, video or binary blob.

\subsubsection{Project}
A project is what is created to contain all content and containers related to a real project. Files can be changed 
within a project and the system can contain several projects and their virtual content are completely disjoint.

\subsubsection{Container}
A container is a virtual folder within a project which can contain content and other containers.

\subsubsection{Snapshot}
A snapshot is a read-only container from the state which the container the was in when the snapshot was created. 
A snapshot can not be updated and can only be deleted from the root of the snapshot. Snapshots are by default stored 
as siblings to the container which they were made from, but they can be contained by any container.

\subsubsection{File}
A file refers to an actual physical file. Files are stored in the database to make backup, deployment and migration easier.

@@

\subsection{JPF}
Java path finder was used to show that the model and plan of how to build the system was sound. The model was built in Java 
with the objective of being as reduced and simple as possible, without loosing any of the cases that needed to be covered 
by the model checker. As the users are mainly going to be handled by external systems they were not included in the model.

Each collection in the persistent storage was emulated by using the built-in ConcurrentHashMap type. 
Each client was represented by a thread and each action taken by the client was randomised. The id hashes which MongoDB is 
using for each entity was imported from the mongo-java-driver-2.13.3 and each object had its own id, generated in the same 
fashion as the real implementation is using, randomly generated by the ObjectId class to minimise collisions that is. 
Furthermore no locking or transactions were used and the threads were running fully concurrently, without any sleep statements.

@@

\newpage
\begin{lstlisting}[label=JPFRESULT,caption=Results of JPF run]
elapsed time:       14:26:53
states:             new=160853259,
                    visited=451102505,
                    backtracked=611955764,
                    end=21640
search:             maxDepth=380,
                    constraints=0
choice generators:  thread=160853255 
                    (signal=0,
                    lock=3603938,
                    sharedRef=146989208,
                    threadApi=3,
                    reschedule=10260106), 
                    data=0

@@

\newpage 
\section{Copy-on-Write}
As the persistent storage does not implement transactions or locks a lot of different problems can occur when several 
clients are working on the same data set at the same time. Such problems could be race conditions and determining the 
happened-before relation. In this work this problem is solved by implementing Copy-on-Write.

\subsection{Comparison of Copy-on-Write system implementations}
\subsubsection{BTRFS}
Btrfs is a B-tree file system for Linux which makes use of Copy-on-Write to make it able to do efficient writeable 
snapshots and clones. It also supports cloning of subtrees without having to actually copy the whole subtree, 
this is due to the Copy-on-Write effect. As several nodes in the tree can refer to the same node each node keeps 
track of how many parents it has by a reference counter so that the node can be deallocated once the node does not 
have any parents any more. The reference counter is not stored in the nodes themselves but rather in a separate data 
structure so that a nodes counter can be modified without modifying the node itself and therefore eludes the 
Copy-on-Write that would have to occur.

\begin{figure}[htp] \centering{
    \includegraphics[scale=0.4]{newtree.png}}
    \caption{Cloning mechanism of Btrfs~\cite{BTRFS}}
    \label{fig:btrfs_tree}
\end{figure}

\subsubsection{Mach kernel}
In the mid 80s when the development of the Mach kernel started, there was problems with that physically copying memory 
was too slow. Too minimise the copying of memory, Copy-on-Write was implemented. It was implemented so that virtual 
copy operations could be done and so that tasks could share read-write memory~\cite{MACH}.

TODO: Insert more systems here

\subsubsection{Perius}
Perius is the implementation that was done to solve the problem at hand at Uprise. This implementation is far from as 
efficient as the others in most aspects, but more efficient in some. As the implementation is built upon MongoDB as 
persistent storage and not a pure tree structure, single nodes can be fetched in O(1) but when querying for subtrees 
they need to be built first, which takes O(log(n)), where n is the number of nodes in the subtree.

In Perius snapshots and clones are not taken in the fashion which Btrfs uses, which can be seen in 
Figure~\ref{fig:btrfs_tree}. As Perius does not have the tree structure pre-built and each node is instead stored in a 
flat storage space, such operation would be too computationally expensive as trees would have to be merged when 
collisions occur, due to the non-blocking nature of the application. Instead this implementation makes a full copy of 
the meta-data of the tree, but still refers to the same binary files until they are changed, which results in the 
creation of a new node.

@@

\section{Resulting system}
\subsection{Persistent storage}
\subsubsection{MongoDB}
MongoDB was chosen as the persistent storage because of its quick lookups and because of its internal storage 
format called BSON, which is very similar to JSON which the API is using. As the formats are similar, the process of 
marshalling and unmarshalling becomes quite easy between the core code, MongoDB instance and REST interface. 
The second reason was that if the system needs to scale in the future it is very easy to distribute MongoDB and if 
needed the system can easily be migrated to Reactive Mongo, which is an asynchronous and 
non-blocking driver for MongoDB and can therefore make the system scale even further~\cite{REACTIVEMONGO}.

\subsection{API}
REST was chosen as only basic CRUD operations needs to be performed and because the BSON format which is used in 
MongoDB is almost identical ~\cite{BSON} to the standardised JSON format which is used by RESTful services~\cite{JSON}.

\subsubsection{REST Endpoints}
For the front-end to communicate with the back-end, a RESTful service is implemented.
The following endpoints were configured:

\begin{itemize}
  \item projects
      \subitem GET - list all projects
      \subitem POST - create new project
  \item projects/\{id\}
      \subitem GET - get specific project
      \subitem PUT - update existing project
      \subitem DELETE - delete existing project
  \item projects/\{id\}/content
      \subitem POST - create new content in a specific project 
  \item projects/\{id\}/content/\{id\}
      \subitem GET - get specific content in a specific project
      \subitem PUT - update existing content in a specific project
      \subitem DELETE - delete existing content in a specific project

  \item projects/\{id\}/snapshots
      \subitem POST - create new snapshot in a specific project

  \item projects/\{id\}/containers
      \subitem POST - create new container in a specific project 
  \item projects/\{id\}/containers/\{id\}
      \subitem GET - get specific container in a specific project
      \subitem PUT - update existing container
      \subitem DELETE - delete existing container
  \item projects/\{id\}/containers/\{id\}/content
      \subitem POST - create new content in a specific container 
  \item projects/\{id\}/containers/\{id\}/content/\{id\}
      \subitem GET - get specific content in a specific container
      \subitem PUT - update existing content in a specific container
      \subitem DELETE - delete existing content in a specific container
\end{itemize}

As can be seen several expected endpoints are missing, this is intentional as the operations missing can be performed in 
a more efficient way. Such endpoint is for example $GET projects/\{id\}/containers$ as all containers exist in 
$GET projects/\{id\}$ and the interface should present a file structure where both content and containers are shown.

@@
